---
title: "psychencode_metadata_analyses"
author: "Daniel Kiss"
date: "9/26/2023"
output: html_document
---

The following code should be run based on the variables generated in lines 1-72 of psychencode_figures_cleaned.Rmd

Plot CTPs vs. ageOnset for each cohort
```{r}
#Note that CMC does not have ageOnset info - therefore we filter for dataset = GVEX and LIBD
ctp_vs_ageonset = all_estimations_long %>%
  filter(primaryDiagnosis %in% c('Schizophrenia'), cell_type %in% c("Inh_SST", "Inh_PVALB", "Inh_VIP"), dataset %in% c("GVEX", "LIBD")) %>%
  ggplot(aes(x = ageOnset, y = rel_prop, color = dataset)) +
  geom_point(alpha = 0.5, size = 0.5) + 
  geom_smooth(se = F, method = 'lm', fullrange = T) +
  ylab('Relative Cell Prop.') + 
  xlab('Age of Onset') + 
  theme(
    axis.text.x = element_text(size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(vjust = -2, size = 16),
    axis.title.y = element_text(size = 16),
    strip.text.x = element_text(size = 16)) +
  facet_grid(~cell_type) +
  scale_y_continuous(limits = c(-2, 2))

ctp_vs_ageonset

#Now let's plot the difference between ageDeath and ageOnset to calculate diagnosisLength
ctp_vs_diagnosislength = all_estimations_long %>%
  filter(primaryDiagnosis %in% c('Schizophrenia'), cell_type %in% c("Inh_SST", "Inh_PVALB", "Inh_VIP"), dataset %in% c("GVEX", "LIBD")) %>%
  mutate(diagnosisLength = ageDeath - ageOnset) %>%
  ggplot(aes(x = diagnosisLength, y = rel_prop, color = dataset)) +
  geom_point(alpha = 0.5, size = 0.5) + 
  geom_smooth(se = F, method = 'lm', fullrange = T) +
  ylab('Relative Cell Prop.') + 
  xlab('Length of Diagnosis') + 
  theme(
    axis.text.x = element_text(size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(vjust = -2, size = 16),
    axis.title.y = element_text(size = 16),
    strip.text.x = element_text(size = 16)) +
  facet_grid(~cell_type) +
  scale_y_continuous(limits = c(-2, 2))

ctp_vs_diagnosislength

#ageOnset vs. ageDeath
ageonset_vs_agedeath = all_estimations %>%
  filter(primaryDiagnosis %in% c('Schizophrenia'), dataset %in% c("GVEX", "LIBD")) %>%
  ggplot(aes(x = ageOnset, y = ageDeath, color = dataset)) +
  geom_point(alpha = 0.5, size = 0.5) + 
  geom_smooth(se = F, method = 'lm', fullrange = T) +
  ylab('Age of Death.') + 
  xlab('Age of Onset') + 
  theme(
    axis.text.x = element_text(size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(vjust = -2, size = 16),
    axis.title.y = element_text(size = 16),
    strip.text.x = element_text(size = 16))
  scale_y_continuous(limits = c(-2, 2))

ageonset_vs_agedeath
                                
```

Replot ageOnset vs. CTP but regress out ageDeath
```{r, fig.width=3, fig.height=2}
#Model for ageOnset
ageOnset_model = all_estimations_long %>% 
  filter(primaryDiagnosis %in% c('Schizophrenia'), cell_type %in% c("Inh_SST", "Inh_PVALB", "Inh_VIP"), dataset %in% c("GVEX", "LIBD")) %>% 
  group_by(dataset, cell_type) %>%
    do(tidy(lm(scale(rel_prop) ~ scale(ageDeath) + scale(RIN) + scale(PMI) + reportedGender +  scale(ageOnset), data = .)))

CELL_TYPES = c("Inh_SST", "Inh_PVALB", "Inh_VIP")
DATASETS = c("GVEX", "LIBD")
ageOnset_residuals_data = matrix(ncol = (ncol(all_estimations_long)+1), nrow = 0) %>% data.frame()
colnames(ageOnset_residuals_data) = c(colnames(all_estimations_long), "resid")

for(DATASET in DATASETS) {
  for(CELL_TYPE in CELL_TYPES) {
    #Create res_model without ageOnset variable
    res_model_data = all_estimations_long %>% filter(primaryDiagnosis %in% c('Schizophrenia'), cell_type == CELL_TYPE , dataset == DATASET)
    #Model without ageOnset
    res_model = lm(scale(rel_prop) ~ scale(ageDeath) + scale(RIN) + scale(PMI) + reportedGender, data = res_model_data)
    #Calculate residuals and add column to res_model_data
    res_model_data = res_model_data %>% add_residuals(var = "resid", model = res_model)
    ageOnset_residuals_data = ageOnset_residuals_data %>% rbind(res_model_data)
  }
}

#Plot ageOnst vs. Residuals
resid_vs_ageonset = residuals_data %>%
  ggplot(aes(x = ageOnset, y = resid, color = dataset)) +
  geom_point(alpha = 0.5, size = 0.5) + 
  geom_smooth(se = F, method = 'lm', fullrange = T) +
  ylab('Residualized CTP Estimate') + 
  xlab('Age of Onset') + 
  theme(
    axis.text.x = element_text(size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(vjust = -2, size = 16),
    axis.title.y = element_text(size = 16),
    strip.text.x = element_text(size = 16)) +
  facet_grid(cols = vars(cell_type), rows = vars(dataset)) +
  scale_y_continuous(limits = c(-2, 2)) + 
  scale_color_manual(values = c("dodgerblue2", "firebrick2"))

resid_vs_ageonset

```

Let's look into correlations between cause of death and the residuals of PV and SST proportions.
```{r}
CELL_TYPES = c("Inh_SST", "Inh_PVALB", "Inh_VIP")
DATASETS = c("GVEX", "LIBD")
residuals_data = matrix(ncol = (ncol(all_estimations_long)+1), nrow = 0) %>% data.frame()
colnames(residuals_data) = c(colnames(all_estimations_long), "resid")

for(DATASET in DATASETS) {
  for(CELL_TYPE in CELL_TYPES) {
    #Create res_model 
    res_model_data = all_estimations_long %>% filter(primaryDiagnosis %in% c('Schizophrenia', 'control'), cell_type == CELL_TYPE , dataset == DATASET)
    res_model = lm(scale(rel_prop) ~ scale(ageDeath) + scale(RIN) + scale(PMI) + reportedGender, data = res_model_data)
    #Calculate residuals and add column to res_model_data
    res_model_data = res_model_data %>% add_residuals(var = "resid", model = res_model)
    residuals_data = residuals_data %>% rbind(res_model_data)
  }
}

#Add filtered causeDeath column
residuals_data  = residuals_data %>%
    mutate(causeDeath_filtered = ifelse(grepl("SUIC", toupper(causeDeath)), "suicide", "other"))

#Plot causeDeath vs. Residuals
resid_vs_ageonset = residuals_data %>%
  filter(primaryDiagnosis == 'Schizophrenia') %>%
  ggplot(aes(x = causeDeath_filtered, y = ageDeath)) +
  geom_boxplot() +
  ylab('Age of Death') + 
  xlab('Cause of Death') + 
  theme(
    axis.text.x = element_text(size = 13, angle = 90),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(vjust = -2, size = 16),
    axis.title.y = element_text(size = 16),
    strip.text.x = element_text(size = 16)) +
  facet_grid(cols = vars(newStudy)) 
  #scale_y_continuous(limits = c(-2, 2)) 
resid_vs_ageonset

#Plot causeDeath vs. Residuals
resid_vs_ageonset = residuals_data %>% 
  filter(primaryDiagnosis == 'Schizophrenia') %>%
  ggplot(aes(x = causeDeath_filtered, y = resid, color = newStudy)) +
  geom_boxplot() +
  ylab('Residualized CTP Estimate') + 
  xlab('Cause of Death') + 
  theme(
    axis.text.x = element_text(size = 13, angle = 90),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(vjust = -2, size = 16),
    axis.title.y = element_text(size = 16),
    strip.text.x = element_text(size = 16)) +
  facet_grid(cols = vars(cell_type), rows = vars(newStudy)) 
  #scale_y_continuous(limits = c(-2, 2)) 
resid_vs_ageonset

```


Let's make some more figures. Here's what we need to illustrate:
-Having SCZ when dying younger = less interneurons
-Dying by suicide = less interneurons 
-Dying by suicide and dying young w/ SCZ are almost completely confounded 
-Having SCZ and dying younger = more brain pathology
-Having SCZ and dying older = paradoxically less brain pathology 

Plot 1:
Re-do figure 2 from the main figures, but with only GVEX and LIBD 
-Model using controls and SCZ + Suicide deaths
-Model using controls and SCZ - Suicide deaths

```{r}
# Create function to map age to age range
age_to_range <- function(age) {
  if (age >= 90) {
    "90+"
  } else {
    lower_bound <- floor(age / 10) * 10
    upper_bound <- lower_bound + 9
    paste0(lower_bound, "-", upper_bound)
  }
}
#Create aggregated causeDeath column
all_estimations_long$age_range = sapply(all_estimations_long$ageDeath, age_to_range)
all_estimations_long = all_estimations_long %>%
    mutate(causeDeath_filtered = ifelse(grepl("SUIC", toupper(causeDeath)), "suicide",
                                        ifelse(grepl("DROWNING|STRANGULATION|INFECTION|ASPHYX|FALL|ACCIDENT|TRAUMA|MVA|OD|HOMIC|HEMORR", toupper(causeDeath)), "accident",
                                        ifelse(grepl("CARD|PULM", toupper(causeDeath)), "cardio",
                                        ifelse(grepl("NATURAL|CANCER|CIRROSIS|COPD|PNEU|SLEEP|PANCREAT|ASTHM", toupper(causeDeath)), "natural",
                                               "other")))))
```

```{r}
#Filter to include all controls, and only SCZ cases that did NOT die by suicide -> Cut off at ageDeath < 90 to have enough data for grouping
mega_lms_by_age_no_suicide = all_estimations_long %>% 
  filter(dataset %in% c("GVEX", "LIBD"), ageDeath < 90) %>% 
  filter(primaryDiagnosis == 'control' | (primaryDiagnosis == 'Schizophrenia' & causeDeath_filtered == "other")) %>%
  group_by(age_range, cell_type) %>%
  do(tidy(lm(scale(rel_prop) ~ scale(RIN) + scale(PMI) + reportedGender + primaryDiagnosis + newStudy, data = .))) %>%
  ungroup() %>% 
  mutate(padj = p.adjust(`p.value`, method = 'BH')) %>%
  mutate(class = case_when(
    str_detect(cell_type, "Inh") ~ "Inhibitory",
    str_detect(cell_type, "Exc") ~ "Excitatory",
    TRUE ~ "Non-Neuronal")) %>%
  mutate(term = recode(term,
                       `(Intercept)` = "Intercept",
                       `reportedGendermale` = "gender:Male",
                       `primaryDiagnosisSchizophrenia` = "SCZ",
                       `scale(ageDeath)` = "ageDeath",
                       `scale(PMI)` = "PMI", 
                       `newStudyNIMH_HBCC` = "NIMH_HBCC",
                       `newStudyPitt` = "Pitt",
                       `newStudyMSSM` = "MSSM",
                       `newStudyPenn` = "Penn",
                       `scale(RIN)` = "RIN"))


#Plot it
mega_lms_by_age_no_suicide %>% 
  filter(term %in% 'SCZ' & cell_type %in% c("Inh_SST", "Inh_PVALB", "Inh_VIP")) %>% 
  mutate(cell_type = fct_reorder(cell_type, estimate)) %>% 
  ggplot(aes(x = age_range, y = estimate)) + 
  geom_hline(yintercept = 0) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error)) + 
  ylab('Beta Coefficient') + 
  xlab('\n \n Age Range') +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 13), 
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(vjust = 8, size = 16),
    axis.title.y = element_text(size = 16),
    strip.text.x = element_text(size = 16)) +
  facet_wrap(~cell_type, drop = T, scale = "free_x") +
  guides(fill = "none") +
  scale_y_continuous(limits = c(-1.0, 1.4))

#This looks weird -> likely because removing CMC data removes a LOT of older subjects

```

In ageDeath <50, SCZ, and interneurons:
-Barplot CTP for suicide deaths vs. accident vs. cardiovascular 
-Scatterplot for CTP vs ageDeath in only suicide deaths 
```{r}
CELL_TYPES = c("Inh_SST", "Inh_PVALB", "Inh_VIP")
DATASETS = c("GVEX", "LIBD")


residuals_data = matrix(ncol = (ncol(all_estimations_long)+1), nrow = 0) %>% data.frame()
colnames(residuals_data) = c(colnames(all_estimations_long), "resid")
for(DATASET in DATASETS) {
  for(CELL_TYPE in CELL_TYPES) {
    #Create res_model -> no ageDeath
    res_model_data = all_estimations_long %>% filter(cell_type == CELL_TYPE , dataset == DATASET, primaryDiagnosis %in% c("control", "Schizophrenia"))
    res_model_data$primaryDiagnosis = res_model_data$primaryDiagnosis %>% factor(levels = c("control", "Schizophrenia"))
    res_model = lm(scale(rel_prop) ~ primaryDiagnosis + scale(ageDeath) + scale(RIN) + scale(PMI) + reportedGender, data = res_model_data)
    #Calculate residuals and add column to res_model_data
    res_model_data = res_model_data %>% add_residuals(var = "resid", model = res_model)
    residuals_data = residuals_data %>% rbind(res_model_data)
  }
}

residuals_data %>% filter() %>%
  ggplot(aes(x = causeDeath_filtered, y = resid)) +
  geom_boxplot(outlier.shape = NA, show.legend = F, notch = F, aes(fill = causeDeath_filtered, alpha = 0.5)) +
  geom_beeswarm(size = 3, alpha = 0.3, aes(colour = causeDeath_filtered), show.legend = F) +
  facet_grid(cols = vars(cell_type)) +
  geom_signif(comparisons = combn(c("suicide", "accident", "natural", "cardio"), 2) %>% t() %>% split(1:4), map_signif_level = TRUE)



all_estimations_long %>%
mutate(causeDeath_filtered = ifelse(grepl("SUIC", toupper(causeDeath)), "suicide", "other")) %>%
 filter(primaryDiagnosis %in% c('Schizophrenia', 'control'), 
        cell_type %in% c("Inh_SST", "Inh_PVALB", "Inh_VIP"), 
        dataset %in% c("GVEX", "LIBD")) %>%
  ggplot(aes(x = ageDeath, y = rel_prop, color = causeDeath_filtered)) +
  geom_point(size = 0.5) +
  geom_smooth(method = lm, se = F) +
  scale_y_continuous(limits = c(-2.5, 2.5)) +
  facet_grid(cols = vars(primaryDiagnosis), rows = vars(dataset))
  
```

What have we learned?
-It's basically impossible to disentangle young ageDeath from death by suicide 
-AgeDeath and CauseDeath are related, 
# Assign the value 10 to the variable x
x <- 10

# Create a sequence from 1 to 5 with an interval of 1
y <- seq(1, 5, by = 1) 

# Add x to each element in y and store the result in z
z <- x + y

# Print the value of z
print(z)# This code generates a sequence of numbers from 1 to 10 using the colon operator
numbers <- 1:10

# Print the sequence of numbers
numbers# This code demonstrates some basic operations in R

# Create a variable with a numeric value
number <- 10

# Create a variable with a character value
word <- "Hello"

# Print the value of the number variable
print(number)

# Print the value of the word variable
print(word)

# Perform addition and store the result in a new variable
sum_result <- number + 5

# Perform concatenation and store the result in a new variable
concat_result <- paste(word, "World")

# Print the sum_result
print(sum_result)

# Print the concat_result
print(concat_result)# This code snippet is a simple R program that prints "Hello World!" to the console

# Output the text "Hello World!" to the console
print("Hello World!")# This code reads a CSV file containing data and performs some basic data manipulation operations in R.

# Import the required library
library(dplyr)

# Read the CSV file into a data frame called 'data'
data <- read.csv("data.csv")

# View the structure of the data frame
str(data)

# Select specific columns from the data frame
selected_columns <- data %>% select(col1, col2)

# Filter the data based on a condition
filtered_data <- selected_columns %>% filter(col1 > 10)

# Sort the data frame based on a column
sorted_data <- filtered_data %>% arrange(col1)

# Perform other data manipulation operations as needed

# Write the final data frame to a new CSV file
write.csv(sorted_data, "output.csv", row.names = FALSE)# This code is written in R programming language and is intended to perform some calculations on a dataset.

# Import the necessary packages for data manipulation and analysis
library(dplyr)

# Read the dataset from a file into a variable called "data"
data <- read.csv("mydata.csv")

# Remove any missing values from the dataset
data <- na.omit(data)

# Perform some calculations on the dataset
# Calculate the mean of a variable called "x" and store it in a new variable called "mean_x"
mean_x <- mean(data$x)

# Calculate the sum of a variable called "y" and store it in a new variable called "sum_y"
sum_y <- sum(data$y)

# Print the calculated mean and sum
print(mean_x)
print(sum_y)# This is a simple R script that calculates the sum of two numbers

# Assign the values of 4 and 6 to the variables 'a' and 'b', respectively
a <- 4
b <- 6

# Calculate the sum of 'a' and 'b' and assign it to the variable 'sum_result'
sum_result <- a + b

# Print the result to the console
print(sum_result)```R
# This code calculates the sum of two numbers
# Declare two variables and assign values
x <- 5
y <- 10

# Calculate the sum of the two numbers
sum <- x + y

# Print the sum
print(sum)
```

\# This code calculates the sum of two numbers
\# Declare two variables and assign values
\# The variable `x` is assigned the value 5
\# The variable `y` is assigned the value 10
`x <- 5`
`y <- 10`

\# Calculate the sum of the two numbers
\# The sum of `x` and `y` is assigned to the variable `sum`
`sum <- x + y`

\# Print the sum
\# The value of `sum` is printed to the console
`print(sum)`# Creating a numeric vector with values 1, 2, 3, 4, 5
vec <- c(1, 2, 3, 4, 5)

# Adding 10 to each element of the vector
vec <- vec + 10

# Printing the updated vector
print(vec)

# Output:
# [1] 11 12 13 14 15# This is a comment line. It is used to provide explanations or context about the code.

# This line of code assigns the value 5 to the variable "x"
x <- 5

# This line of code assigns the value 10 to the variable "y"
y <- 10

# This line of code calculates the sum of variables "x" and "y" and assigns the result to the variable "sum"
sum <- x + y

# This line of code prints the value of the variable "sum" to the console
print(sum)
```R
# This code calculates the mean of a numeric vector

# Create a numeric vector
numbers <- c(1, 2, 3, 4, 5)

# Calculate the mean of the vector
mean_numbers <- mean(numbers)

# Print the mean
print(mean_numbers)
```













