---
title: "psychencode_supplemental_figures"
author: "Daniel Kiss"
date: "3/7/2024"
output: html_document
---

Packages etc.
```{r}
library(dplyr)
library(tidyverse)
library(edgeR)
library(markerGeneProfile) 
library(matrixStats)
library(cowplot)
library(broom)
library(knitr)
library(ggpubr)
library(biomaRt)
library(ggrepel)
library(patchwork)
library(ggsignif)
library(modelr)
library(ggbeeswarm)
library(lemon)
theme_set(theme_classic2())
#Colour palette
cbPalette = c("#E69F00", "#D55E00", "#CC79A7")
```

Loading Bulk Data
```{r}

#Add PMI info from CMC metadata
psychencode_metadata = read.csv(("/external/rprshnas01/external_data/psychencode/PsychENCODE/Metadata/CapstoneCollection_Metadata_Clinical.csv"))
CMC_metadata = read.csv("/external/rprshnas01/external_data/psychencode/PsychENCODE/CMC/Metadata/SYNAPSE_TABLE_QUERY_123020650.csv") %>% filter(dataType == "geneExpression", fileFormat == "tsv")  %>% left_join(psychencode_metadata)
names(CMC_metadata)[names(CMC_metadata) == "Individual_ID"] = "individualID"

#Cell type proportions for each data set merged with metadata - ready for modelling 
acc_estimations = read_csv("/external/rprshnas01/kcni/dkiss/cell_prop_psychiatry/data/cmc_acc_mgp_estimations.csv") %>% filter(primaryDiagnosis %in% c("control", "Schizophrenia")) %>% left_join(CMC_metadata[, c("individualID", "PMI")], by = "individualID")
acc_estimations$newStudy = acc_estimations$newStudy %>% factor(levels = c("Pitt_ACC", "MSSM_ACC", "Penn_ACC"))
acc_estimations_long = acc_estimations %>% pivot_longer(cols = Astrocyte:VLMC, names_to = 'cell_type', values_to = 'rel_prop') %>%
  mutate(cell_class = case_when(
    cell_type %in% c("LAMP5", "PAX6", "PVALB", "SST", "VIP") ~ "Inhibitory",
    cell_type %in% c("IT", "L4.IT", "L5.ET", "L5.6.IT.Car3", "L5.6.NP", "L6.CT", "L6b") ~ "Excitatory",
    TRUE ~ "Non-Neuronal"
  ))

```

Cohort demographic summary
```{r}
acc_summary_df = acc_estimations %>%
    distinct(individualID, .keep_all = TRUE) %>% # Ensure rows are unique based on individualID
    group_by(newStudy) %>%
    summarise(
        CohortName = first(newStudy),
        PercentMale = mean(reportedGender == "male", na.rm = TRUE) * 100,
        AvgAge = mean(ageDeath, na.rm = TRUE),
        AgeSD = sd(ageDeath, na.rm = TRUE),
        NumCases = sum(primaryDiagnosis == "Schizophrenia", na.rm = TRUE),
        NumControls = sum(primaryDiagnosis == "control", na.rm = TRUE)
    )
```

Loading LCM proportions
```{r}
#Load Sibille Lab microscopy
sibille_lab_cell_props = sibille_lab_cell_fractions = read_csv('/external/rprshnas01/netdata_kcni/stlab/cross_cohort_MGPs/sibille_lab_cell_fractions.csv') %>% filter(Subject.Group %in% c("Control", "SCHIZ")) %>% distinct()

# this is the tetrad metadata
tetrad_path = '/external/rprshnas01/netdata_kcni/stlab/Collaborators/Sibille/RNAseq_processed/'
tetrad_meta = read_csv(paste0(tetrad_path, "pTable with correct med info.csv"))
tetrad_meta$sample_names = str_sub(tetrad_meta$X, end=-5) %>% make.names()
tetrad_meta$sample_names =  gsub("_", ".", tetrad_meta$sample_names)
tetrad_meta = tetrad_meta %>% mutate(Subject = HU., 
                       Reported.Gender = case_when(Sex == "M" ~"Male",
                                                   Sex == "F" ~ "Female"),
                       Age_norm = Age, 
                       PMI..in.hours. = PMI) %>% filter(Cell.Type == "SST")

microCTPs = left_join(sibille_lab_cell_props, tetrad_meta) %>% filter(Subject != 1367)
microCTPs_long =  microCTPs %>% pivot_longer(cols = c(PVALB, PYR_23, PYR_56, SST, VIP ), names_to = 'cell_type', values_to = 'rel_prop') 

```

Load LCM DE table
```{r}
scz_lcm_seq_de = readxl::read_excel("~/cell_prop_psychiatry/data/psychencode_lcm_DE_scz.xlsx")
```


Model rCTPs based on diagnosis and covariates separately in each cohort
```{r}
acc_combined_lms = acc_estimations_long %>%
  group_by(newStudy, cell_type) %>%
  do(tidy(lm(scale(rel_prop) ~ scale(RIN) + scale(ageDeath) + scale(PMI) +
               reportedGender + primaryDiagnosis, data = .))) %>%
  ungroup() %>%
  mutate(padj = p.adjust(p.value, method = 'fdr'),
         term = recode(term,
                       `(Intercept)` = "Intercept",
                       `reportedGendermale` = "gender:Male",
                       `primaryDiagnosisSchizophrenia` = "SCZ",
                       `scale(ageDeath)` = "Age",
                       `scale(PMI)` = "PMI",
                       `scale(RIN)` = "RIN")) %>%
  merge(unique(acc_estimations_long[c('dataset', 'newStudy')]), by.x = "newStudy") %>%
  merge(unique(acc_estimations_long[c('cell_class', 'cell_type')]), by.x = "cell_type") 

acc_combined_lms$newStudy = acc_combined_lms$newStudy %>% factor(levels = c("Pitt_ACC", "MSSM_ACC", "Penn_ACC"))
```

Interaction mega-analysis w/ all data 
```{r}
acc_mega_lms = acc_estimations_long %>% 
  group_by(cell_type) %>%
  do(tidy(lm(scale(rel_prop) ~ scale(RIN) + scale(ageDeath) + scale(PMI) +
             reportedGender + primaryDiagnosis + newStudy + scale(ageDeath)*primaryDiagnosis, data = .))) %>%
  ungroup() %>% 
  mutate(padj = p.adjust(`p.value`, method = 'fdr'),
         term = recode(term,
                       `(Intercept)` = "Intercept",
                       `reportedGendermale` = "gender:Male",
                       `primaryDiagnosisSchizophrenia` = "SCZ",

                       `scale(ageDeath)` = "Age",
                       `scale(ageDeath):primaryDiagnosisSchizophrenia` = "Age x Diagnosis",
                       `scale(PMI)` = "PMI", 
                       `newStudyNIMH_HBCC` = "NIMH_HBCC",
                       `newStudyLIBD_szControl` = "LIBD_szControl",
                       `newStudyPitt` = "Pitt_ACC",
                       `newStudyMSSM` = "MSSM_ACC",
                       `newStudyPenn` = "Penn_ACC",
                       `scale(RIN)` = "RIN")) %>%
  merge(unique(acc_estimations_long[c('cell_class', 'cell_type')]), by.x = "cell_type") 
```

Functions to  generate age range bins 
```{r}
# Create function to map age to age range
age_to_range = function(age) {
    if(age >= 15 & age < 20) {
    return("15-19")
  } else {
    lower_bound = floor((age - 20) / 10) * 10 + 20
    upper_bound = lower_bound + 9
    return(paste0(lower_bound, "-", upper_bound))
  }
}
# Create function to bin ages to >=70 and <70
classify_ages = function(ages) {
  classifications = ifelse(ages <= 70, "Under 70", "Over 70")
  return(classifications)
}

# Create function to map age to age range based on 20% quantiles
age_to_range_quantile <- function(age) {
  quantiles = quantile(age, probs = seq(0.5, 1, 0.5))
  range_vec = c()
  for (i in 1:length(quantiles)) {
    if (i == 1) {
      range_vec[age <= quantiles[i]] <- paste0(0, "-", quantiles[i])
    } else {
      range_vec[age > quantiles[i-1] & age <= quantiles[i]] <- paste0(quantiles[i-1], "-", quantiles[i])
    }
  }
  range_vec
}
```

Create age groups from age 20-90 in increments of 10 years. Plot meganalysis beta coefs against age group.
```{r}
acc_estimations_long$age_range = sapply(acc_estimations_long$ageDeath, age_to_range)

acc_lms_by_age = acc_estimations_long %>% filter(ageDeath >= 30) %>%
  group_by(age_range, cell_type) %>%
  do(tidy(lm(scale(rel_prop) ~ scale(RIN) + scale(PMI) +
             reportedGender + primaryDiagnosis + newStudy, data = .))) %>%
  ungroup() %>% 
  mutate(padj = p.adjust(`p.value`, method = 'fdr'),
         term = recode(term,
                       `(Intercept)` = "Intercept",
                       `reportedGendermale` = "gender:Male",
                       `primaryDiagnosisSchizophrenia` = "SCZ",
                       `scale(ageDeath)` = "Age",
                       `scale(PMI)` = "PMI", 
                       `newStudyNIMH_HBCC` = "NIMH_HBCC",
                       `newStudyPitt` = "Pitt_ACC",
                       `newStudyMSSM` = "MSSM_ACC",
                       `newStudyPenn` = "Penn_ACC",
                       `scale(RIN)` = "RIN")) %>%
  merge(unique(acc_estimations_long[c('cell_class', 'cell_type')]), by.x = "cell_type") 
```



Bulk RNAseq Meganalysis based on age +/- 70
```{r}
acc_estimations_long$age_class <- sapply(acc_estimations_long$ageDeath, classify_ages) %>% factor(levels = c("Under 70", "Over 70"))
acc_mega_lms_by_age_class = acc_estimations_long %>% 
  group_by(age_class, cell_type) %>%
  do(tidy(lm(scale(rel_prop) ~ scale(RIN) + scale(PMI) +
             reportedGender + newStudy + primaryDiagnosis, data = .))) %>%
  ungroup() %>% 
  mutate(padj = p.adjust(`p.value`, method = 'fdr'),
         term = recode(term,
                       `(Intercept)` = "Intercept",
                       `reportedGendermale` = "gender:Male",
                       `primaryDiagnosisSchizophrenia` = "SCZ",
                       `scale(ageDeath)` = "Age",
                       `scale(PMI)` = "PMI", 
                       `newStudyNIMH_HBCC` = "NIMH_HBCC",
                       `newStudyPitt` = "Pitt",
                       `newStudyMSSM` = "MSSM",
                       `newStudyPenn` = "Penn",
                       `scale(RIN)` = "RIN")) %>%
  merge(unique(acc_estimations_long[c('cell_class', 'cell_type')]), by.x = "cell_type")


```

Model Microscopy CTPs
```{r}

micro_mega_lms = microCTPs_long %>% 
  group_by(cell_type) %>%
  do(tidy(lm(scale(rel_prop) ~ scale(PMI) + Subject.Group + Sex + scale(Age),  data = .))) %>%
  ungroup() %>% 
  mutate(padj = p.adjust(`p.value`, method = 'fdr')) %>%
   mutate(term = recode(term,
                       `(Intercept)` = "Intercept",
                       `scale(PMI)` = "PMI", 
                       `Subject.GroupControl` = "Control",
                       `Subject.GroupSCHIZ`= "SCZ",
                       `SexM` = "Sex"))

```



------------////FIGURES////------------

Figure S1: Illustrating the association between cell prop and diagnosis and cohort specific differences
```{r, fig.height=8, fig.width=8 }
significant_combined = subset(acc_combined_lms, padj < 0.1 & cell_type %in% c("SST", "PVALB", "VIP") & term == "SCZ")
significant_combined$asterisks <- ifelse(significant_combined$padj <= 0.01, "***",
                                         ifelse(significant_combined$padj <= 0.05, "**", "*"))
significant_combined$vjust <- ifelse(significant_combined$estimate >= 0, -3, 4)

figure_S1b = acc_combined_lms %>%
  filter(term == 'SCZ', cell_type %in% c("SST", "PVALB", "VIP")) %>% 
  mutate(cell_type = fct_reorder(cell_type, estimate)) %>% 
  ggplot(aes(x = newStudy, y = estimate, fill = newStudy)) + 
  geom_hline(yintercept = 0) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error)) + 
  ylab('Beta Coefficient') + 
  xlab('') +
  scale_fill_manual(values = cbPalette) +
  geom_text(data = significant_combined,
            aes(newStudy, estimate, label = asterisks), 
            vjust = significant_combined$vjust) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(vjust = 6, size = 15),
    axis.title.y = element_text(size = 16),
    strip.text.x = element_text(size = 14),
    axis.line.x = element_blank(), 
    axis.ticks.x = element_blank()) +
  facet_grid(~cell_type, drop = T, scale = "free_x", space = "free") +
  guides(fill = guide_none()) +
  scale_y_continuous(limits = c(-1, 1))

# Figure 1c: Raw MGPs vs. Diagnosis
figure_S1c <- acc_estimations_long %>%
  filter(cell_type %in% c("SST", "PVALB", "VIP")) %>%
  ggplot(aes(x = primaryDiagnosis, y = rel_prop)) +
  geom_boxplot(outlier.shape = NA, aes(fill = newStudy, alpha = 0.5), show.legend = F, notch = T) +
  geom_beeswarm(size = 1, alpha = 0.3, aes(colour = newStudy), show.legend = F) +
  ylab('Relative Cell Prop.') + 
  xlab('Diagnosis') + 
  scale_x_discrete(labels = c("C", "SCZ")) +
  scale_colour_manual(values = cbPalette) +
  scale_fill_manual(values = cbPalette) +
  theme(
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(vjust = -2, size = 15),
    axis.title.y = element_text(size = 15),
    strip.text.x = element_text(size = 14),
    strip.text.y = element_text(size = 14)) +
  guides(fill = FALSE) +
  facet_grid(rows = vars(cell_type), cols = vars(newStudy), drop = T, scale = "free", switch = "y", space = "free") +
  geom_signif(comparisons = list(c("control", "Schizophrenia")), map_signif_level = TRUE, , test = wilcox.test) +
  scale_y_continuous(limits = c(-4, 4)) 

#Figure 1d - Hisotgram of cohort ages
figure_S1d = acc_estimations %>%
    ggplot(aes(x = ageDeath)) +
    geom_histogram(aes(fill = newStudy)) +
    facet_wrap(~ newStudy, nrow = 6, scales = "free_y") +
    labs(x = "ageDeath", y = "Frequency") +
    scale_fill_manual(values = c("#E69F00", "#D55E00", "#CC79A7")) + 
  scale_y_continuous(n.breaks = 4) +
  guides(fill = guide_none()) + 
  theme(panel.grid = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size = rel(0.9), face = "plain"),
    axis.text.x = element_text(size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text( size = 15),
    axis.title.y = element_text(size = 16),
    strip.text.x = element_text(size = 14))

# Combine Figure 1
figure_S1_top = (figure_S1b | figure_S1d) + plot_layout(widths = c(2,1))
figure_S1_bottom = figure_S1c
  
figure_S1 = figure_S1_top / figure_S1_bottom + plot_annotation(tag_levels = 'A') + plot_layout(heights = c(2, 3)) & theme(plot.tag = element_text(size = 25))
figure_S1

#Save to plots folder
output_filename <- "~/cell_prop_psychiatry/plots/figure_S1.jpg"

# Use ggsave to save the plot as a JPG image
ggsave(output_filename, figure_S1, width = 10, height = 10, dpi = 200)

```



Figure 2: Introducing and presenting an interaction between age and diagnosis w/rt cell proportion.
```{r, fig.height=10, fig.width=10}
# Plot interaction beta values from meganalysis
significant_mega = subset(acc_mega_lms, padj <= 0.1 & term == "Age x Diagnosis")
significant_mega$asterisks = ifelse(significant_mega$padj <= 0.01, "***",
                                   ifelse(significant_mega$padj <= 0.05, "**", "*"))
significant_mega$vjust = ifelse(significant_mega$estimate >= 0, -2, 3)

# Figure 2a: Results from interaction meganalysis stratified by cell class - plot interaction beta coeff
figure_S2a = acc_mega_lms %>% 
  filter(term %in% 'Age x Diagnosis') %>% 
  ggplot(aes(x = cell_type, y = estimate)) + 
  geom_hline(yintercept = 0) + 
  geom_bar(stat = "identity", position = "dodge") + 
  facet_grid(~cell_class, drop = T, scale = "free_x", space = "free") +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error)) + 
  ylab('Interaction Coefficient') + 
  xlab('Cell Type') + 
  geom_text(data = significant_mega,
            aes(cell_type, estimate, label = asterisks), 
            vjust = significant_mega$vjust) +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 12), 
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(vjust = 8, size = 15),
    axis.title.y = element_text(size = 15),
    strip.text.x = element_text(size = 14)) +
  guides(fill = FALSE) +
  scale_y_continuous(limits = c(-0.5, 0.7))

# Figure 2b: Plot beta values from meganalysis in 10-year age bins for SST, PVALB, VIP
significant_mega_age = subset(acc_lms_by_age, padj <= 0.1 & term == "SCZ" & cell_type %in% c("SST", "PVALB", "VIP"))
significant_mega_age$asterisks = ifelse(significant_mega_age$padj <= 0.01, "***",
                                         ifelse(significant_mega_age$padj <= 0.05, "**", "*"))
significant_mega_age$vjust = ifelse(significant_mega_age$estimate >= 0, -3, 3)

figure_S2b = acc_lms_by_age %>% 
  filter(term %in% 'SCZ' & cell_type %in% c("SST", "PVALB", "VIP")) %>% 
  mutate(cell_type = fct_reorder(cell_type, estimate)) %>% 
  ggplot(aes(x = age_range, y = estimate)) + 
  geom_hline(yintercept = 0) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error)) + 
  ylab('Beta Coefficient') + 
  xlab('\n \n Age Range') + 
  geom_text(data = significant_mega_age,
            aes(age_range, estimate, label = asterisks), 
            vjust = significant_mega_age$vjust) +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 13), 
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(vjust = 8, size = 15),
    axis.title.y = element_text(size = 15),
    strip.text.x = element_text(size = 14)) +
  facet_wrap(~cell_type, drop = T, scale = "free_x") +
  guides(fill = FALSE) +
  scale_y_continuous(limits = c(-1.4, 1.4))

# Add residuals to all_estimations_long
res_model = lm(scale(rel_prop) ~ scale(RIN) + scale(PMI) + reportedGender , data = acc_estimations_long)
acc_estimations_long = acc_estimations_long %>% add_residuals(var = "resid", model = res_model)

# Figure 2c: Visualizing interaction between age and diagnosis by plotting rel_prop vs. age in each diagnosis 
figure_S2c = acc_estimations_long %>%
  filter(cell_type %in% c("SST", "PVALB", "VIP")) %>%
  ggplot(aes(x = ageDeath, y = rel_prop, color = primaryDiagnosis)) +
  geom_point(alpha = 0.5, size = 0.5) + 
  geom_smooth(se = F, method = 'lm', fullrange = T) +
  ylab('Relative Cell Prop.') + 
  xlab('Age') + 
  theme(
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(vjust = -1, size = 15),
    axis.title.y = element_text(size = 15),
    legend.position = c(0.9, 0.8),
    legend.title = element_blank(),
    legend.text = element_text(size = 12),  
    strip.text.x = element_text(size = 14)) +
  facet_grid(~cell_type, drop = T, scale = "free_x", space = "free") +
  guides(fill = guide_legend(nrow = 1)) +
  scale_y_continuous(limits = c(-2, 2)) + 
  scale_color_manual(values = c("dodgerblue2", "firebrick2"), labels=c('CON', 'SCZ'))

#Figure 2d: Mega-analysis results when binned as +/- 70 years
significant_mega_age_class = subset(acc_mega_lms_by_age_class, padj <= 0.1 & term == "SCZ" & cell_type %in% c("SST", "PVALB", "VIP"))
significant_mega_age_class$asterisks = ifelse(significant_mega_age_class$padj <= 0.01, "***", ifelse(significant_mega_age_class$padj <= 0.05, "**", "*"))
significant_mega_age_class$vjust = ifelse(significant_mega_age_class$estimate >= 0, -2, 3)

figure_S2d = acc_mega_lms_by_age_class %>% 
  filter(term %in% 'SCZ' & cell_type %in% c("SST", "PVALB", "VIP")) %>% 
  mutate(cell_type = fct_reorder(cell_type, estimate)) %>% 
  ggplot(aes(x = age_class, y = estimate)) + 
  geom_hline(yintercept = 0) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error)) + 
  ylab('Beta Coefficient') + 
  xlab('\n \n Age Range') + 
  geom_text(data = significant_mega_age_class,
            aes(age_class, estimate, label = asterisks), 
            vjust = significant_mega_age_class$vjust) +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 13), 
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(vjust = 8, size = 15),
    axis.title.y = element_text(size = 15),
    strip.text.x = element_text(size = 14)) +
  facet_wrap(~cell_type, drop = T, scale = "free_x") +
  guides(fill = FALSE) +
  scale_y_continuous(limits = c(-1, 1))

figure_S2_top = (figure_S2b +figure_S2d) + plot_layout(widths = c(1.6,1))
figure_S2 = figure_S2_top / figure_S2c / figure_S2a
figure_S2 = figure_S2 + plot_annotation(tag_levels = 'A') + plot_layout(heights = c(1, 1, 1)) & theme(plot.tag = element_text(size = 25))
figure_S2

#Save to plots folder
output_filename <- "~/cell_prop_psychiatry/plots/figure_S2.jpg"

# Use ggsave to save the plot as a JPG image
ggsave(output_filename, figure_S2, width = 11, height = 11, dpi = 200)

```

Figure S3: Validating trends and interactions in ssRNAseq 
```{r, fig.height= 5, fig.width=9}
# Figure 3a: Relative cell proportions from snCTPs
figure_S3a <- microCTPs_long %>% 
  filter(Subject.Group %in% c('SCHIZ', "Control"), cell_type %in% c("SST", "PVALB", "VIP")) %>%
  ggplot(aes(x = Age_norm, y = rel_prop, color = Subject.Group)) +
  geom_point(alpha = 0.8, size = 1.3) + 
  geom_smooth(se = F, method = 'lm') +
  scale_colour_hue(labels = c("Control", "SCZ")) +
  ylab('Relative Cell Prop.') + 
  xlab('Age') + 
  theme(
    axis.text.x = element_text(size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(vjust = -2, size = 16),
    axis.title.y = element_text(size = 16),
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 12), 
    strip.text.x = element_text(size = 15),
    strip.text.y = element_text(size = 15)
  ) +
  scale_color_manual(values = c("dodgerblue2", "firebrick2"), labels=c('CON', 'SCZ')) +
  facet_wrap(~cell_type, drop = T, scale = "free")

# Figure 3b: microCTP Beta plot
significant_micro_mega <- subset(micro_mega_lms, padj <= 0.1 & term == "SCZ" & cell_type %in% c("SST", "PVALB", "VIP"))
significant_micro_mega$asterisks <- ifelse(significant_micro_mega$padj <= 0.01, "***",
                                         ifelse(significant_micro_mega$padj <= 0.05, "**", "*"))
significant_micro_mega$vjust <- ifelse(significant_micro_mega$estimate >= 0, -5, 5)

figure_S3b <- micro_mega_lms %>% 
  filter(term %in% 'SCZ' & cell_type %in% c("SST", "PVALB", "VIP")) %>% 
  mutate(cell_type = fct_reorder(cell_type, estimate)) %>% 
  ggplot(aes(x = cell_type, y = estimate)) + 
  geom_hline(yintercept = 0) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error)) + 
  ylab('Beta Coefficient') + 
  xlab('Cell Type') + 
  geom_text(data = significant_micro_mega,
            aes(estimate, label = asterisks), 
            vjust = significant_micro_mega$vjust) +
  theme(
    axis.text.x = element_text(size = 13),
    axis.text.y = element_text(size = 13),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 12), 
    strip.text.x = element_text(size = 15)
  ) 



# Figure 3: Combined figure
figure_S3 = (figure_S3a + figure_S3b) +
  theme(plot.margin = unit(c(5.5, 10, 5.5, 5.5), "pt"))
figure_S3 = figure_S3 + plot_annotation(tag_levels = 'A') + plot_layout(widths = c(3, 1)) & theme(plot.tag = element_text(size = 25))
figure_S3

#Save to plots folder
output_filename <- "~/cell_prop_psychiatry/plots/figure_S3.jpg"

# Use ggsave to save the plot as a JPG image
ggsave(output_filename, figure_S3, width = 9, height = 4, dpi = 200)

```

```{r}
MGPVolcano <- function(matrix, marker_genes) {
  significant_genes <- subset(matrix, abs(logFC) > 1 & P.Value < 0.1)
  significant_marker_genes <- subset(matrix, gene_symbol %in% marker_genes & abs(logFC) > 1 & P.Value < 0.1)
  
  # Increase plot limits slightly to provide more space for labels
  x_range <- range(matrix$logFC, na.rm = TRUE)
  y_range <- range(-log10(matrix$P.Value), na.rm = TRUE)
  
  matrix %>%
    ggplot(aes(x = logFC, y = -log10(P.Value))) + 
    geom_vline(xintercept = -1, linetype = "dotted") +
    geom_vline(xintercept = 1, linetype = "dotted") +
    geom_hline(yintercept = -log10(0.1), linetype = "dotted") +
    geom_point(
      aes(label = gene_symbol), 
      alpha = 0.3, # Set opacity 30%
      color = "black", # Default color for non-significant points
      size = 3 # Adjust size as needed
    ) +
    geom_point(
      data = significant_marker_genes, 
      color = 'blue', # Set specific color for significant marker genes
      size = 3 # Keep size consistent
    ) +
    geom_point(
      data = subset(matrix, gene_symbol %in% marker_genes & !(gene_symbol %in% significant_marker_genes$gene_symbol)),
      color = 'red', # Set specific color for other specified marker genes
      size = 3 # Keep size consistent
    ) +
    geom_label_repel(
      data = significant_marker_genes, 
      aes(label = gene_symbol), 
      box.padding = 0.35, 
      point.padding = 0.5,
      segment.color = 'grey50',
      fill = "white", 
      color = "blue",
      max.overlaps = Inf,
      nudge_x = (x_range[2] - x_range[1]) * 0.2, # Nudge labels horizontally
      nudge_y = (y_range[2] - y_range[1]) * 0.2 # Nudge labels vertically
    ) +
    geom_label_repel(
      data = subset(matrix, gene_symbol %in% marker_genes & !(gene_symbol %in% significant_marker_genes$gene_symbol)), 
      aes(label = gene_symbol), 
      box.padding = 0.35, 
      point.padding = 0.5,
      segment.color = 'grey50',
      fill = "white", 
      color = "red",
 max.overlaps = Inf,
      nudge_x = (x_range[2] - x_range[1]) * 0.1, # Nudge labels horizontally
      nudge_y = (y_range[2] - y_range[1]) * 0.1 # Nudge labels vertically
    ) +
    theme_cowplot() +
    theme(legend.position = "none") +
    expand_limits(x = x_range + c(-1, 2) * 0.5, y = y_range + c(-1, 1) * 0.5) # Expand plot limits
}


```
Figure S4: Volcano plots for LCM-seq 
```{r, fig.width= 12, fig.height=4}
markers = read.csv("~/cell_prop_psychiatry/data/MGP_markers.csv")
SST_markers = markers %>% filter(subclass == "SST") %>% pull(gene)
PVALB_markers = markers %>% filter(subclass == "PVALB") %>% pull(gene)
VIP_markers = markers %>% filter(subclass == "VIP") %>% pull(gene)

scz_lcm_seq_de = scz_lcm_seq_de %>% filter(-log10(P.Value) < 5)


#Set figure font sizes
common_theme = theme(
  axis.title = element_text(size = 16),
  axis.text = element_text(size = 16)
)

SST_volcano = scz_lcm_seq_de %>% filter(CT == "SST") %>% MGPVolcano(SST_markers) + ggtitle("SST") +
                    theme(plot.title = element_text(hjust = 0.5, size = 18)) + common_theme
PVALB_volcano = scz_lcm_seq_de %>% filter(CT == "PVALB") %>% MGPVolcano(PVALB_markers) + ggtitle("PVALB") +
                    theme(plot.title = element_text(hjust = 0.5, size = 18)) + common_theme
VIP_volcano = scz_lcm_seq_de %>% filter(CT == "VIP") %>% MGPVolcano(VIP_markers) + ggtitle("VIP") +
                    theme(plot.title = element_text(hjust = 0.5, size = 18)) + common_theme

figure_S4 = ggarrange(SST_volcano, PVALB_volcano, VIP_volcano,
                      labels = c("C", "D", "E"),
                      ncol = 3, nrow = 1,
                      font.label = list(size = 32, face = "bold"))

#Save to plots folder
output_filename <- "~/cell_prop_psychiatry/plots/figure_S4.jpg"

# Use ggsave to save the plot as a JPG image
ggsave(output_filename, figure_S4, width = 12, height = 4, dpi = 200)
```

