---
title: "psychencode_DE_gprofiler"
author: "Daniel Kiss"
date: "8/17/2023"
output: html_document
---

As a preliminary GO/pathway analysis, we will use G:Profiler to look at differentially expressed pathways in our candidate cell types. We will begin by generating txt file containing each cell type's DE genes in both Above_70 and Below_70 groups to be used in gprofiler.

```{r}
DE_results_MtSinai = readRDS("/external/rprshnas01/kcni/dkiss/cell_prop_psychiatry/data/DE_results_MtSinai")
DE_results_McLean = readRDS("/external/rprshnas01/kcni/dkiss/cell_prop_psychiatry/data/DE_results_McLean")
CELL_TYPES = names(DE_results_McLean) 

top_genes_count <- 2000

for (CELL_TYPE in CELL_TYPES) {
  ranked_genes_under = DE_results_McLean[[CELL_TYPE]]$DE_tables$DE_under %>%
    arrange(desc(logFC)) %>%
    pull(genes) %>%
    head(top_genes_count)
  ranked_genes_over = DE_results_McLean[[CELL_TYPE]]$DE_tables$DE_over %>%
    arrange(desc(logFC)) %>%
    pull(genes) %>%
    head(top_genes_count)
  
  file_path = paste0("/external/rprshnas01/kcni/dkiss/cell_prop_psychiatry/data/gprofiler_files/", CELL_TYPE, "_ranked_genes.txt")
  
  # Create a connection to the file
  con <- file(file_path, "w")
  
  # Write the under_70 list
  cat(paste0(">",CELL_TYPE, "_under_70\n", paste(ranked_genes_under, collapse = "\n")), file = con, sep = "\n")
  
  # Write the over_70 list
  cat(paste0(">",CELL_TYPE, "_over_70\n", paste(ranked_genes_over, collapse = "\n")), file = con, sep = "\n")
  
  # Close the file connection
  close(con)
}

```


Keon's Code for running ClusterProfiler in R -> THIS WILL NOT RUN IN R < 4.1***
```{r}
#Because this won't run in R > 4.1, we'll run it on a different device and upload results to the SCC as object GO_plot_list - it also takes a long time to run UPDATE: We can run this in bash if you update R there

library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)

DE_results_McLean = readRDS("DE_results_McLean")
DE_results_MtSinai = readRDS("DE_results_MtSinai")

CELL_TYPES = names(DE_results_McLean)

GO_plot_list = list()

for(CELL_TYPE in CELL_TYPES) {
  GENE_LIST_UNDER = DE_results_McLean[[CELL_TYPE]]$DE_tables$DE_under %>%
  arrange(desc(logFC)) %>% pull(logFC, genes)
  GENE_LIST_OVER = DE_results_McLean[[CELL_TYPE]]$DE_tables$DE_over %>%
  arrange(desc(logFC)) %>% pull(logFC, genes)
  
gseGO_under = clusterProfiler::gseGO(geneList = GENE_LIST_UNDER, 
  ont = "BP", 
  keyType = "SYMBOL", 
  minGSSize = 15, 
  maxGSSize = 500, 
  pvalueCutoff = 0.05, 
  verbose = TRUE, 
  OrgDb = org.Hs.eg.db, 
  pAdjustMethod = "none")
gseGO_under_s = simplify(gseGO_under, cutoff = 0.7, by = "p.adjust", select_fun = min)
res = gseGO_under_s@result
GO_plot_under = enrichplot::dotplot(gseGO_under_s, showCategory = 15) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 50)) +
  facet_grid(~.sign) 

gseGO_over = clusterProfiler::gseGO(geneList = GENE_LIST_OVER, 
  ont = "BP", 
  keyType = "SYMBOL", 
  minGSSize = 15, 
  maxGSSize = 500, 
  pvalueCutoff = 0.05, 
  verbose = TRUE, 
  OrgDb = org.Hs.eg.db, 
  pAdjustMethod = "none")
gseGO_over_s = simplify(gseGO_over, cutoff = 0.7, by = "p.adjust", select_fun = min)
res = gseGO_over_s@result
GO_plot_over = enrichplot::dotplot(gseGO_over_s, showCategory = 15) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 50)) +
  facet_grid(~.sign) 

GO_plot_combined = GO_plot_under + GO_plot_over

GO_plot_list[[CELL_TYPE]] = GO_plot_combined
}

write_rds(GO_plot_list, "/external/rprshnas01/kcni/dkiss/cell_prop_psychiatry/data/McLean_GSEA_Plots")

```






